---
title: "Student Ethogram - Data Analysis"
author: "Marinara Marcato"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DescTools)
library(ggplot2)
library(dplyr)
library(plyr)
```

## Introduction 
This document shows the data analysis carried out to investigate 
dog's behavioral data considering the ethograms filled out by students
Hazel and Conghal.

## Data Exploration

### Columns: Variables
Importing the ethograms for analysis.
```{r, echo = FALSE}
setwd("C://Users//marinara.marcato//Data//Ethogram//Researchers")
dc1 <- read.csv('dc1.csv', stringsAsFactors=FALSE)
dc2 <- read.csv('dc2.csv', stringsAsFactors=FALSE)
dcs <- read.csv('dcs.csv', stringsAsFactors=FALSE)
```

The ethogram comprised of the following types and number of variables.
```{r, echo = FALSE}
# separate columns by data types
ordinal = c(
  'Familiarisation.Response..Oriented.to.Handler.',
  'Familiarisation.Response..Exploration.',
  'Familiarisation.Response..Waiting.',
  'Call.Back.Response',
  'Walking.Distractibility',
  'Walking.Pull.on.leash',
  'Walking.Pull.strength',
  'Walking.Initiative',
  'Sitting.Attempts',
  'Sitting.Response',
  'Lying.Attempts',
  'Lying.Response',
  'Lying.Settled',
  'Body.check.Response',
  'Distractions.First.Response..Object.',
  'Distractions.First.Response..Human.',
  'Distractions.First.Response..Car.',
  'Distractions.First.Response..Food.',
  'Distractions.Second.Response..Object.',
  'Distractions.Second.Response..Human.',
  'Distractions.Second.Response..Car.',
  'Distractions.Second.Response..Food.',
  'Distractions.Pull.on.leash',
  'Distractions.Pull.strength',
  'Kong.Concentration.Response',
  'Kong.Retrieve.Response.to.stimulus',
  'Kong.Retrieve.Response.to.assessor',
  'Kong.Retrieve.Back',
  'Dog.Response',
  'Dog.Call.Back.Response',
  'Crate.Entering',
  'Crate.Behaviours..Relaxing.',
  'Crate.Behaviours..Sniffing.Exploring..',
  'Crate.Behaviours..Attention.Seeking.',
  'Crate.Behaviours..Digging.',
  'Crate.Behaviours..Licking.Mouthing.',
  'Crate.Behaviours..Whining.',
  'Crate.Behaviours..Barking.',
  'Crate.Response')

count = c(
  'Short.Barks.Count',
  'Continuous.Barks.Duration',
  'Continuous.Barks.Count',
  'Whines.Duration',
  'Whines.Count',
  'Shakes.Count',
  'Jumps.Count')

categorical = c(
  'Body.check.General..Mouths.', 
  'Body.check.General..Licks.',
  'Tea.Towel.First.Response..Indifferent.', 
  'Tea.Towel.First.Response..Turns.head.',
  'Tea.Towel.First.Response..Attempts.to.Removes.towel.by.moving.',
  'Tea.Towel.First.Response..Attempts.to.Removes.towel.with.mouth.',
  'Tea.Towel.First.Response..Plays.',
  'Tea.towel.Second.Response..Indifferent.',
  'Tea.towel.Second.Response..Turns.head.',
  'Tea.towel.Second.Response..Attempts.to.Removes.towel.by.moving.',
  'Tea.towel.Second.Response..Attempts.to.Removes.towel.with.mouth.',
  'Tea.towel.Second.Response..Plays.',
  'Kong.Retrieve.Lateralisation'
  # eliminate 'Crate.Urinating' because only has No
)

all = c(ordinal, count, categorical)

cat("Ordinal:", length(ordinal))       # 39 variables
cat("Count:", length(count))         # 7 variables
cat("Categorical:", length(categorical))   # 13 variables
cat("All:", length(all))           # 59 variables in total
```

### Rows: Dogs
Total number of dogs per Data Collection:
```{r , echo = FALSE}
# printing out information
cat("DC1:", length(dc1$Name), "\n", paste(sort(dc1$Name), collapse = ","))
# printing out information
cat("DC2:", length(dc2$Name), "\n", paste(sort(dc2$Name), collapse = ","))
# removing litter with only one dog
dcs_litter = subset(dcs,  !Litter %in%  c('L','M','T', 'Z'))
cat("DCS:", length(unique(dcs_litter$Name)), "\n", paste(sort(unique(dcs$Name)), collapse = ","))
```

Plotting some demographics
```{r , echo = FALSE}
prop_sex = rbind(table(dc1$Sex),table(dc2$Sex),table(dcs$Sex)/2)
barplot(prop_sex, beside = TRUE, space = c(0.1, 0.5), ylim = c(0,(1.2*max(prop_sex))), col = rainbow(nrow(prop_sex)), 
main = " Number of Dogs per Gender considering Data Collections")

legend("topright", legend = c("DC1", "DC2", "DC1&2"), col = rainbow(nrow(prop_sex)), pch = 16)

prop_litter = rbind.fill(data.frame(rbind(table(dc1$Litter))),data.frame(rbind(table(dc2$Litter))),data.frame(rbind(table(dcs$Litter)/2)))
barplot(as.matrix(prop_litter), beside = TRUE, space = c(0.1, 0.5), ylim = c(0,(1.2*max(prop_litter, na.rm = TRUE))), col = rainbow(nrow(prop_litter)), 
main = "Number of Dogs per Litter considering Data Collections" )
legend("topright", legend = c("DC1", "DC2", "DC1&2"), col = rainbow(nrow(prop_litter)), pch = 16)
```


## Data Analysis
### Between Litters, Controlling for DC
The difference in behaviours between different Litters will be investigated controlling for Data Collection.

##### Ordinal Variables
Kruskal-Wallis Test is used to test whether the difference between different litters is significant. 
Dunn Test using bonferroni is used as a post-hoc to find out which groups differ. 


```{r, echo = FALSE}
# Kruskal-Wallis Test:
# testing significant differences between multiple groups 
ordinal_between_groups <- function(df){
  result = c()
  for (variable in ordinal){
      r = kruskal.test(df[, variable], as.factor(df$Litter))
      result = rbind(result, c(variable, r[1], r[2], r[3]))
  }
  return (result)
}

# saves results for ALL ORDINAL-scored behaviours between litters
dc1_ordinal = ordinal_between_groups(dc1)
dc2_ordinal = ordinal_between_groups(dc2)
# saves results for SIGNIFICANTLY DIFFERENT ORDINAL-scored behaviours between litters 
dc1_ordinal_sign = dc1_ordinal[dc1_ordinal[,4]<0.05, ]
dc2_ordinal_sign = dc2_ordinal[dc2_ordinal[,4]<0.05, ]
```


```{r, echo = FALSE}
# POST-HOC Test for Kruskal-Wallis Test: 
#   pairwise comparison of groups which showed statistically significant difference
posthoc_between_groups <- function(df, variables){
  results = c()
  for (variable in variables){
    r = DunnTest(df[,variable], as.factor(df$Litter), method = "bonferroni")
    results = rbind(results, c(variable,  r))
  }
  return(results)
}
# performs posthoc on SIGNIFICANTLY DIFFERENT ORDINAL variables
dc1_ordinal_ph = posthoc_between_groups(dc1, unlist(dc1_ordinal_sign[,1]))
dc2_ordinal_ph = posthoc_between_groups(dc2, unlist(dc2_ordinal_sign[,1]))
```


```{r, echo = FALSE}
# prints outputs to file
cat('DC 1 - Difference between litters')
print(dc1_ordinal_sign)
cat('DC 2 - Difference between litters')
print(dc2_ordinal_sign)

plot_ordinal <- function(df, var){
  # returns bar plot of @var in df per 'Litter'
  df = df %>% group_by(Litter) %>% summarise_at(var, funs(mean))
  barplot(df[[var]], names.arg = df[['Litter']], ylab = 'Mean score', xlab= 'Litter', main = var)
}

for (var in unlist(dc1_ordinal_sign[,1])){
  print (var)
  plot_ordinal(dc1, var)   
}
```

#### Count Variables

No need to perform Post Hoc because none of the variables was significantly different 

```{r, echo = TRUE}
count_between_groups <- function(df){
  result = c()
  for (variable in count){
    r = kruskal.test(df[, variable], df$Litter)
    #print(c(variable, r[3]))
    result = rbind(result, c(variable, r[1], r[2], r[3]))
  }
  return (result)
}
dc1_count = count_between_groups(dc1)
dc2_count = count_between_groups(dc2)
dc1_count_sign = dc1_count[dc1_count[,4]<0.05, ]
dc2_count_sign = dc2_count[dc2_count[,4]<0.05, ]
```

#### Categorical Variables

Testing significance of all CATEGORICAL variables considering GROUPS
```{r, echo = TRUE}

categorical_between_groups <- function(df){
  result = c()
  for (variable in categorical){
    if (length(unique(df[,variable]))>1){
      r = chisq.test(df[, variable], df$Litter, simulate.p.value=TRUE)
      #print(c(variable, r[3]))
      result = rbind(result, c(variable, r[1], r[2], r[3]))
    }
  }
  return (result)
}

dc1_categorical = categorical_between_groups(dc1)
dc2_categorical = categorical_between_groups(dc2)
dc1_categorical_sign = dc1_categorical[dc1_categorical[,4]<0.05, ]
dc2_categorical_sign = dc2_categorical[dc2_categorical[,4]<0.05, ]
print(dc1_categorical_sign)
print(dc2_categorical_sign)
```

#### Visualising overall resposes in DC1 and DC2
```{r, echo = TRUE}
# barplot of behaviour per DC
for (variable in ordinal){
  par(mfrow = c(1, 3))
  for(i in unique(dcs$Data.Collection.Number)) {
    barplot(table(dcs[dcs$Data.Collection.Number == i, variable]), space = 1,
         xlab = "",
         col = rainbow(length(unique(dcs[dcs$Data.Collection.Number == i, variable]))),
         main = paste("DC", i, variable)) 
  }
  boxplot(dcs[dcs$Data.Collection.Number == 1,variable], dcs[dcs$Data.Collection.Number == 2,variable], names = unique(dcs$Data.Collection.Number))
  #title(strwrap(descriptive, width=60), cex.main=1, font.main=1)
}
```
