---
title: "Trainers Ethogram - Outcome (GD, AD, W) - Post-Hoc"
author: "Marinara Marcato"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/marinara.marcato/Project/Scripts/dog_ethogram")
knitr::opts_chunk$set(echo = TRUE)
#install.packages("ggpubr")
library(DescTools)
library(ggplot2)
library(dplyr)
library(plyr)
library(tidyverse)
library(rstatix)
library(ggpubr)
```

## Introduction 
This document shows the data analysis carried out to investigate 
dog's behavioral data considering the ethograms filled out by two trainers.

## Data Exploration
### Columns: Variables
The ethogram comprised of the following types and number of variables.

```{r, echo = FALSE}
dc1 <- read.csv('0_data//2_analysis//2021-05-03_Ethogram-Trainers-DC1.csv', stringsAsFactors=FALSE)
dc2 <- read.csv('0_data//2_analysis//2021-05-03_Ethogram-Trainers-DC2.csv', stringsAsFactors=FALSE)

vars <- read.csv('0_data//3_results//trainers_outcome.csv', stringsAsFactors=FALSE)
```


### Rows: Dogs
Total number of dogs with DC1 and DC2:
```{r, echo = FALSE}
## dogs with dc1
cat('Dogs with DC1: ', length(dc1$Name))
# removing dogs in Training
dc1_outcome = dc1 %>% filter( Status != 'in Training' )
# printing out information
cat("DC1 & Outcome:", length(dc1_outcome$Name)) # "\n" paste(sort(dc1_outcome$Name), collapse = ", "))

## dogs with dc2 
cat('Dogs with DC2: ', length(dc2$Name))
# removing dogs in Training
dc2_outcome = dc2 %>% filter( Status != 'in Training' )
# printing out information
cat("DC2 & Outcome:", length(dc2_outcome$Name)) # "\n", paste(sort(dc2_outcome$Name), collapse = ", "))
```

## Data Analysis: Post-Hoc
TBD
Testing significance of all ORDINAL variables considering Training Outcome as grouping factor.
It saves results for significantly different behaviours between litters for each dc.
```{r, eval  = FALSE, echo = FALSE}
ordinal_between_groups <- function(var, data, group){
    print(var)
    print(group)
    print(length(data[,var]))
    print(length(data[, group]))
    #r = kruskal.test(df[, var], df[,group])
    #result = c(r[1], r[2], r[3])
    #return (result)
}
#vars_dc1_ordinal =  vars  %>% select('Variable')
vars = vars %>% filter( dc1_sign == 1 , Type == 'ordinal' ) %>% 
        mutate(dc1_outcome = mapply(FUN = ordinal_between_groups, var = vars$Variable, MoreArgs = list(dc1_outcome, 'Status')))

#dc1_ordinal = count_between_groups(dc1_outcome)
v_ft = vars %>% filter( dc1_sign == 1 , Type == 'ordinal' ) %>% pull('Variable')
```

## Outputs: Statistically Significant Variables

### DC1
Below are the variables that showed statistically significant differences in behaviours at DC1 considering between different Training Outcomes ('AD', 'GD', 'W').
```{r, echo  = FALSE}
vars_dc1 =  vars %>% filter( dc1_sign == 1 ) %>% pull('Variable')
print(vars_dc1)

plotseries <- function(var){
    ggplot(dc1_outcome, aes_string( x = var, fill = 'Status')) + 
            geom_bar(position = "dodge")}
lapply(vars_dc1, plotseries )
```

### DC2
Below are the variables that showed statistically significant differences in behaviours at DC2 considering different Training Outcomes ('AD', 'GD', 'W').

```{r,echo = FALSE}
vars_dc2 =  vars %>% filter( dc2_sign == 1 ) %>% pull('Variable')
print(vars_dc2)
plotseries <- function(var){
    ggplot(dc2_outcome, aes_string( x = var, fill = 'Status')) + 
            geom_bar(position = "dodge")}
lapply(vars_dc2, plotseries )
```
