---
title: "Student Ethogram - Data Analysis"
author: "Marinara Marcato"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DescTools)
library(ggplot2)
library(dplyr)
library(plyr)
```

## Introduction 
This document shows the data analysis carried out to investigate 
dog's behavioral data considering the ethograms filled out by students
Hazel and Conghal.


## Data Exploration

### Columns: Variables
The ethogram comprised of the following types and number of variables.

```{r, echo = FALSE}
setwd("C://Users//marinara.marcato//Data//Ethogram//")
data <- read.csv('20-02-26_Ethogram.csv')
vars <- read.csv('variables.csv')

# Adding a column to df vars for Test 
vars = vars %>% mutate(Test = case_when(
        Type == "count" ~ "Wilcoxon",
        Type == "ordinal" ~ "Wilcoxon", 
        Type == "categorical" ~ "Chisq"))

# print the number of variables per data type
print(table(vars$Type))
```

### Rows: Dogs
Total number of ethograms in the dataset, there is two per dog:
```{r data, echo = FALSE}
# removing duplicates, 105 examples left
data = data[!duplicated(data[,3:4]),]
# data collection number per dog
dog_dc = aggregate(data$Data.Collection.Number, by = list(Category = data$Name), FUN = sum)
# retrieving name of dogs with dc1 and dc2
dogs_dcs = dog_dc[dog_dc$x ==3, 'Category']
# selecting dogs with dc1 and dc2 only
dcs = data %>% filter( Name %in% dogs_dcs)
# ordering the df by Name and DC
dcs = dcs[order(dcs$Name, dcs$Data.Collection.Number),]
dcs$Name = factor(dcs$Name)
cat("Number of dogs with DC1 and DC2: ", length(unique(dcs$Name)))
```


## Within Subject Analysis

result = wilcox.test(dcs[dcs$Data.Collection.Number == 1 , "Isolation.Response..Time.oriented."],  dcs[dcs$Data.Collection.Number == 2, "Isolation.Response..Time.oriented."], paired = TRUE)
class(result[["statistic"]])
, 
```{r, include = TRUE}

wilcox <- function(var){
  #r = wilcox.test(dcs[dcs$Data.Collection.Number == 1 , var],  dcs[dcs$Data.Collection.Number == 2, var], paired = TRUE)
  return(12)
  #return(c(312,12))
}
vars =  vars %>% mutate(Result = if_else(
        Test == "Wilcoxon", wilcox(vars$Variable), 
        ifelse( Test == "Chisq", 422 , NA)))

vars =  vars %>% mutate(Result = case_when(
        Test == "Wilcoxon" ~ wilcox(Variable), 
        Test == "Chisq" ~ 4))

str(dcs[vars[vars$Type == "count","Variable"],])
dcs[vars['Type']== "ordinal"]

```
vars$Result
colnames(vars)
unique(vars$Test)


vars[vars$Test == "Wilcoxon", ]

### Ordinal data
Running Wilcoxon for ordinal data. 
```{r, include = TRUE}
# groups are the dcs
ordinal_within_groups <- function(df){
  result = c()
  for (variable in ordinal){
      r = wilcox.test(df[df$Data.Collection.Number == 1 , variable], df[df$Data.Collection.Number == 2, variable], paired = TRUE)
      result = rbind(result, c(variable, r[1], r[2], r[3]))
  }
  return (result)
}

dcs_ordinal = ordinal_within_groups(dcs)
ordinal_sign = dcs_ordinal[dcs_ordinal[,4]<0.05, ]
print(ordinal_sign)
```

### Count data
Running wilcoxon for count data. No signficant results found.
```{r, include = TRUE}
# groups are the dcs
count_within_groups <- function(df){
  result = c()
  for (variable in count){
      r = wilcox.test(df[df$Data.Collection.Number == 1 , variable], df[df$Data.Collection.Number == 2, variable], paired = TRUE)
      result = rbind(result, c(variable, r[1], r[2], r[3]))
  }
  return (result)
}
dcs_count = count_within_groups(dcs)
count_sign = dcs_count[dcs_count[,4]<0.05, ]
```

### Categorical data
run chi-squared for categorical data

```{r, include = TRUE}
# groups are the dcs
categorical_within_groups <- function(df){
  result = c()
  for (variable in categorical){
    if (length(unique(df[df$Data.Collection.Number == 1 ,variable]))>1 & length(unique(df[df$Data.Collection.Number == 2 ,variable]))>1){
      #`correct=FALSE` to turn off Yates’ continuity correction.? 
      r = chisq.test(df[df$Data.Collection.Number == 1 , variable], df[df$Data.Collection.Number == 2, variable])
      result = rbind(result, c(variable, r[1], r[2], r[3]))
    }
  }
  return (result)
}
dcs_categorical = categorical_within_groups(dcs)
categorical_sign = dcs_categorical[dcs_categorical[,4]<0.05, ]
print(unlist(categorical_sign[,1]))
```
