---
title: "Student Ethogram - Data Analysis"
author: "Marinara Marcato"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DescTools)
library(ggplot2)
library(dplyr)
library(plyr)
```

## Introduction 
This document shows the data analysis carried out to investigate 
dog's behavioral data considering the ethograms filled out by students
Hazel and Conghal.


## Data Exploration

### Columns: Variables
The ethogram comprised of the following types and number of variables.

```{r, echo = FALSE}
setwd("C://Users//marinara.marcato//Data//Ethogram//")
vars <- read.csv('variables.csv')
data <- read.csv('20-02-26_Ethogram.csv')

# Adding a column to df vars for Test 
vars %>% mutate(Test = case_when(
        Type == "count" ~ "Wilcoxon",
        Type == "ordinal" ~ "Wilcoxon", 
        Type == "categorical" ~ "Chisq",
        Type == "info" ~ "NA",
        Type == "unused" ~ "NA"))

# dataframe column names
cols = colnames(data)

print(vars[vars$Type == 'categorical',])
# separate columns by data types
ordinal = c(
  'Familiarisation.Response..Oriented.to.Handler.',
  'Familiarisation.Response..Exploration.',
  'Familiarisation.Response..Waiting.',
  'Call.Back.Response',
  'Walking.Distractibility',
  'Walking.Pull.on.leash',
  'Walking.Pull.strength',
  'Walking.Initiative',
  'Sitting.Attempts',
  'Sitting.Response',
  'Lying.Attempts',
  'Lying.Response',
  'Lying.Settled',
  'Body.check.Response',
  'Distractions.First.Response..Object.',
  'Distractions.First.Response..Human.',
  'Distractions.First.Response..Car.',
  'Distractions.First.Response..Food.',
  'Distractions.Second.Response..Object.',
  'Distractions.Second.Response..Human.',
  'Distractions.Second.Response..Car.',
  'Distractions.Second.Response..Food.',
  'Distractions.Pull.on.leash',
  'Distractions.Pull.strength',
  'Kong.Concentration.Response',
  'Kong.Retrieve.Response.to.stimulus',
  'Kong.Retrieve.Response.to.assessor',
  'Kong.Retrieve.Back',
  'Dog.Response',
  'Dog.Call.Back.Response',
  'Crate.Entering',
  'Crate.Behaviours..Relaxing.',
  'Crate.Behaviours..Sniffing.Exploring..',
  'Crate.Behaviours..Attention.Seeking.',
  'Crate.Behaviours..Digging.',
  'Crate.Behaviours..Licking.Mouthing.',
  'Crate.Behaviours..Whining.',
  'Crate.Behaviours..Barking.',
  'Crate.Response')

count = c(
  'Short.Barks.Count',
  'Continuous.Barks.Duration',
  'Continuous.Barks.Count',
  'Whines.Duration',
  'Whines.Count',
  'Shakes.Count',
  'Jumps.Count')

categorical = c(
  'Body.check.General..Mouths.', 
  'Body.check.General..Licks.',
  'Tea.Towel.First.Response..Indifferent.', 
  'Tea.Towel.First.Response..Turns.head.',
  'Tea.Towel.First.Response..Attempts.to.Removes.towel.by.moving.',
  'Tea.Towel.First.Response..Attempts.to.Removes.towel.with.mouth.',
  'Tea.Towel.First.Response..Plays.',
  'Tea.towel.Second.Response..Indifferent.',
  'Tea.towel.Second.Response..Turns.head.',
  'Tea.towel.Second.Response..Attempts.to.Removes.towel.by.moving.',
  'Tea.towel.Second.Response..Attempts.to.Removes.towel.with.mouth.',
  'Tea.towel.Second.Response..Plays.',
  'Kong.Retrieve.Lateralisation'
  # eliminate 'Crate.Urinating' because only has No
)



all = c(ordinal, count, categorical)

cat("Ordinal:", length(ordinal))       # 39 variables
cat("Count:", length(count))         # 7 variables
cat("Categorical:", length(categorical))   # 13 variables
cat("All:", length(all))           # 60 variables in total
```

### Rows: Dogs
Total number of ethograms in the dataset, there is two per dog:
```{r data, echo = FALSE}
# removing duplicates, 105 examples left
data = data[!duplicated(data[,3:4]),]
# data collection number per dog
dog_dc = aggregate(data$Data.Collection.Number, by = list(Category = data$Name), FUN = sum)
# retrieving name of dogs with dc1 and dc2
dogs_dcs = dog_dc[dog_dc$x ==3, 'Category']
# selecting dogs with dc1 and dc2 only
dcs = data %>% filter( Name %in% dogs_dcs)
# ordering the df by Name and DC
dcs = dcs[order(dcs$Name, dcs$Data.Collection.Number),]
dcs$Name = factor(dcs$Name)
cat(length(dcs))
```


## Within Subject Analysis

### Ordinal data
Running wilcoxon for ordinal data. 
```{r, include = TRUE}
# groups are the dcs
ordinal_within_groups <- function(df){
  result = c()
  for (variable in ordinal){
      r = wilcox.test(df[df$Data.Collection.Number == 1 , variable], df[df$Data.Collection.Number == 2, variable], paired = TRUE)
      result = rbind(result, c(variable, r[1], r[2], r[3]))
  }
  return (result)
}

dcs_ordinal = ordinal_within_groups(dcs)
ordinal_sign = dcs_ordinal[dcs_ordinal[,4]<0.05, ]
print(ordinal_sign)
```

### Count data
Running wilcoxon for count data. No signficant results found.
```{r, include = TRUE}
# groups are the dcs
count_within_groups <- function(df){
  result = c()
  for (variable in count){
      r = wilcox.test(df[df$Data.Collection.Number == 1 , variable], df[df$Data.Collection.Number == 2, variable], paired = TRUE)
      result = rbind(result, c(variable, r[1], r[2], r[3]))
  }
  return (result)
}
dcs_count = count_within_groups(dcs)
count_sign = dcs_count[dcs_count[,4]<0.05, ]
```

### Categorical data
run chi-squared for categorical data
```{r, include = TRUE}
# groups are the dcs
categorical_within_groups <- function(df){
  result = c()
  for (variable in categorical){
    if (length(unique(df[df$Data.Collection.Number == 1 ,variable]))>1 & length(unique(df[df$Data.Collection.Number == 2 ,variable]))>1){
      #`correct=FALSE` to turn off Yates’ continuity correction.? 
      r = chisq.test(df[df$Data.Collection.Number == 1 , variable], df[df$Data.Collection.Number == 2, variable])
      result = rbind(result, c(variable, r[1], r[2], r[3]))
    }
  }
  return (result)
}
dcs_categorical = categorical_within_groups(dcs)
categorical_sign = dcs_categorical[dcs_categorical[,4]<0.05, ]
print(unlist(categorical_sign[,1]))
```
