---
title: "Demographics vs Outcome (GD, AD, W)"
author: "Marinara Marcato"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "C:/Users/marinara.marcato/Project/Scripts/dog_ethogram")
knitr::opts_chunk$set(echo = TRUE)
#reticulate for python engine
#knitr::knit_engines$set(python = reticulate::eng_python)
#library(reticulate)


#install.packages("ggpubr")
library(DescTools)
library(ggplot2)
library(dplyr)
library(plyr)
library(tidyverse)
library(rstatix)
library(ggpubr)
```

## Introduction 
This document tests the hypothesis that there is a difference in demographics for dogs considering their training outcome.

## Data Exploration
### Columns: Variables
The following demographic variables will be considered comprised of the following types and number of variables.

```{r, echo = FALSE}

```


### Rows: Dogs
Total number of dogs analysed:


## Data Analysis
Hypothesis testing: investigate difference in behavioural scores for all variables analysed at (DC1, DC2) considering the training outcome (GD, AD, W). 
The following statistical tests were performed considering the variable type  <br>
Kruskal -> Ordinal = 53 variables in total<br>
Chiqs -> Categorical = 17 variables in total<br>


## Results
Results are saved in a file (trainers_outcome.csv). Below are variable with statistically significant results. <br>
It means that dogs showed a statistically different behaviour taking DC1 at Week 3 and DC2 Week 10. 
